### 2017-03-29 pt_characterization/rep_check/paper.tex
Characterizing drafts

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Repeatability Test}

\author{
Young-Kyoon Suh\\
}
\maketitle

\section{Description}
This document presents two sets of histograms based on the two runs 
of a program under test, called {\em INC}, with a set of increasing task lengths 
from 1 second to 4096 seconds.
We would like to see if both of the sets have the same shape for the same task length.
If so then, we can say that run repeatability is satisfied in our experiment setting.

\section{Experiment Notes}
Table~\ref{tab:exp_notes} provides a short description of our experimental runs, 
on which the following histograms are based.

%% done
\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2cm}|p{3cm}|p{5cm}|p{5cm}|} \hline
Machine & Task Length (sec) & Description & Experiment Period\\ \hline
{\tt sodb9} &  INC1$\sim$INC64 & Two runs with 1000 samples & 2017-03-02 $\sim$ 2017-03-?? / 2017-03-?? $\sim$ 2017-03-??\\ \hline
{\tt sodb9} &  INC128$\sim$INC1024 & Two runs with 300 samples & 2017-03-02 $\sim$ 2017-03-?? / 2017-03-?? $\sim$ 2017-03-??\\ \hline
{\tt sodb10} & INC2048 & Two runs of 300 samples & 2017-03-02 $\sim$ 2017-03-09 / 2017-03-13 $\sim$ 2017-03-20\\ \hline
{\tt sodb12} & INC4096 & Two runs of 300 samples & 2017-02-13 $\sim$ 2017-02-27 / 2017-03-02 $\sim$ 2017-03-17\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Notes on experiment runs used for histograms\label{tab:exp_notes}}
\end{table}

Now we show histograms of elapsed time (ET) and process time (PT) of INC.

\pagebreak

\input{res1.tex}

\pagebreak
\newpage

\input{res2.tex}

\end{document}