### 2017-04-02 sedona-ieice16/final/reviews/review_response.tex
Sedona revision

\documentclass[10pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{algorithmic}
\usepackage{alltt}
\usepackage{array}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
%\usepackage[caption=false]{caption}
\usepackage[dvipsnames]{color}
%\usepackage{colortbl}
%\usepackage{enumitem}
%\usepackage{eqparbox}
%\usepackage{fixltx2e}
%\usepackage{float}
%\usepackage{floatflt}
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%\usepackage{multirow}
\usepackage{stfloats}
%\usepackage[font=footnotesize]{subfig}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\usepackage{subfigure} 
\usepackage{times}
\usepackage{url}
\usepackage{verbatim} 
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{color}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilite{{\bf TODO:} {\em #1}}}

\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}

% new commands: from file -- to permit re-use
%\input{newcommands.tex}

%Hyphenation problems
\hyphenation{name-space}
\hyphenation{schema-location}
\hyphenation{xml-lint}
\hyphenation{time-stamp}
\hyphenation{time-stamps}
\hyphenation{time-stamped}

\urlstyle{rm}
\graphicspath{{./figures/}}

\definecolor{gray}{rgb}{0.5,0.5,0.5}
%\newcommand{\rev}[1]{\textcolor{blue}{#1}}
%\newcommand{\quo}[1]{\textcolor{gray}{#1}}
%\newcommand{\rev}[1]{\vspace{3mm} \noindent {\em {\color{blue}{#1}}}}

\newenvironment{myindentpar}[1]%
{\begin{list}{}
         {\vspace{10pt}
					\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\rev}[1]{\begin{myindentpar}{.25in} {\em {\color{blue}{#1}}}\end{myindentpar}}

\newenvironment{myindentparQ}[1]%
{\begin{list}{}
         {\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\quo}[1]{\begin{myindentparQ}{.25in} {{\color{gray}{#1}}}\end{myindentparQ}}

\begin{document}

\title{Author's Reply}
\author{}
\maketitle

\section*{Overview}\label{sec:overview}
I appreciate your review and address your remaining question below. 
Note that my response to your question is 
incorporated into the revision and marked in {\color{blue}blue}.

\newcounter{RP}

\clearpage
\section*{Reviewer}\label{sec:rev1}

% 1
\rev{
$<<$ Reviewer's comments to the author(s) $>>$

The manuscript is well revised.

However, I have a question how to estimate each daemon's influence as time.
The estimation is important for accuracy of your proposed method.

For acceptance, please clarify the estimation method and its validity.

}

If I interpret your question correctly,
(1)~the identification process of infrequent, long-running daemons 
and (2)~the proposed {\em cutoff} method address your question. 
The cutoff is used as a criterion to determine whether to discard a sample, to improve measurement quality. 
Both are elaborated from Steps~2 through~9 in Figure~3, augmented with a one-line comment (above Step~2), 
for better clarification.

%Here is the summary of how to estimate each daemon's influence as time via the cutoffs. 
To estimate each daemon's influence, we do the following. 
\begin{itemize}
\item We run a program-under-test with 128 seconds, termed PUT128, 800 times 
enough to experience an infrequent, long-running daemon 
that may run every few hours or even once a day (Step~2).

\item From the 800 samples, we render a scatter-plot of pairs of successive samples of elapsed time (ET) of PUT128
as illustrated in Fig.~4~(b) and see if that scatter-plot has an $L$-shape (Step~3).

\item We check if the central cluster emerges (Step~4) by zooming into the scatter-plot (Step~4).

\item For each daemon in the central cluster, we compute the maximum and standard deviation 
of the process time (PT) of that daemon (Step~5). (As specified in the Characteristics paragraph, PT is defined as 
the sum of user and system time, supplied from the kernel via the Netlink interface, before and after each run.)

\item (1)~We identify, for each daemon in the samples in the $L$-shape (or $L$-sample), 
which daemon has its PT greater than two standard deviations above the maximum PT 
of that daemon. (2)~We also collect some extra processes (e.g. {\tt bash}, {\tt rhn\_check}) 
that are not in the cluster but only in the $L$-sample (Step~6). 
The identified processes are treated as infrequent, long-running.

\item For each so-identified infrequent, long-running daemon, we compute the minimum time 
in the $L$-samples (Step~7). (This computation provides the valley between the maximum PT from 
the central cluster and the minimum PT from the $L$-shape.)

\item We repeat Steps 2---6 with the much-longer running of PUT16384, to confirm if 
any of the identified daemons is actually not infrequent at the much larger task length of 16384 seconds (Step~8).
This step is required to prevent the false positive case in our identification process, 
as some daemons with a periodicity of hours are infrequent for a ``short'' program time like 128 seconds.

\item We now compute the {\em cutoff} for each of those identified infrequent, long-running daemons, 
based on the PUT128 and PUT16384 runs. 
\end{itemize}

%(The user- and system-time measures of captured processes are supplied by the kernel, via the Netlink interface.)

For more details, please see the prose in Section~2 marked in blue.

\end{document}