### 2017-04-12 pt_characterization/rep_check/paper.tex
repeatability test draft

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Repeatability Test}

\author{
Young-Kyoon Suh\\
}
\maketitle

\section{Description}
This document presents two sets of histograms based on the two runs 
of a program under test, called {\em INC}, with a set of increasing task lengths 
from 1 second to 4096 seconds.
We would like to see if both of the sets have the same shape for the same task length.
If so then, we can say that run repeatability is satisfied in our experiment setting. 
In this study the used protocol is equivalent to SEDONA, eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5. In other words, the protocol skips the second step 
of EMPv5 that removes samples with process times over/below two standard deviations 
from the average after performing the first step.

Now we show histograms of elapsed time (ET) and process time (PT) of INC via the SEDONA protocol.

\pagebreak

\input{res1.tex}

\pagebreak
\newpage

\input{res2.tex}

\end{document}