### 2017-08-08 subopt-tods17/OLDtex/methodology.tex

\section{Methodology}

\input{model.tex}

\input{detecting.tex}

\input{measuring.tex}

\input{repeatability.tex}

%\input{monotonicity.tex}

\subsection{Summary of Sub-Optimality}
\shorten{
In this paper we utilize a new methodology to ask new
questions about a mature topic, query optimization, in \hbox{order} to develop and
test a novel model and theory about when suboptimality arises, to
arrive at a new understanding of query optimization. This may later enable
new engineering approaches as well as a better understanding of the limits
of existing approaches.

\begin{figure}[bth]\centering
\includegraphics[width=0.50\textwidth]{figures/understanding.eps}
\caption{Empirical generalization ([3, page~5])\label{fig:empirical}}
\end{figure}

This paper utilizes {\em empirical
  generalization}, in which understanding proceeds along two conceptual
dimensions, as shown in Figure~\ref{fig:empirical}. As Cohen~\c2j{\cite{cohenbook}}{[1995]} has
articulated, science (in general, and this paper specifically) progresses
(in the figure, on the 
$y$-axis) from studies of a particular system (e.g, a particular database
query optimization algorithm in the context of a particular database
management system), through studies of a class of systems (e.g.,~a study
{\em across} several disparate DBMSs), to statements about computational
tools in general (e.g.,~a theory that holds for rule-based optimizers,
whether in DBMSs, AI systems, or compilers). This progression increases the
domain of applicability, and thus the generality, of a theory. Science also
progresses (in the figure, on the 
$x$-axis) from description of the phenomenon, to prediction, and eventually
through causal explanation, within an articulated, thoroughly-tested theory.
We seek the rather ambitious outcome of the articulation of a general theory
of suboptimality that applies to the entire class of DBMSes and which
provides causal explanation of suboptimality: when it will arise and why.

Note that this line of study is very different from conventional
approaches summarized above from the literature. We are trying to understand
cost-based query optimizers as a {\em general} class of computational artifacts,
to come up with insights and ultimately with predictive theories about how such
optimizers, again, as a general class, behave~\cite{davies}.}