### 2017-08-08 subopt-tods17/OLDtex/motivation.tex

\section{Motivation}
\todo{Rick has modify token}
Consider a simple select-project-join (SPJ) query, with a few attributes in
the SELECT clause, a few tables referenced in the FROM clause,
and a few equality predicates in the WHERE clause. This might be an
excerpt from a more complex query, with the tables being intermediate
results from other computations. We'll use as an example Query 769 from the
study to be described in Section~\ref{sec:detecting}. This query was one of 1000
queries that were randomly generated.

\vspace*{-2ex}
{\small\begin{verbatim}
        SELECT t2.id3 
        FROM ft_HT1 t3, ft_HT3 t2, ft_HT1 t0, ft_HT1 t1 
        WHERE (t3.id2=t2.id1 AND t2.id1=t0.id4 AND t0.id4=t1.id4)
\end{verbatim}
}

\vspace*{-2ex}
\noindent
This query utilizes four correlation names in the FROM clause, three of
which were associated with the same table, {\tt ft\_HT1}, whose cardinality
statistics were varied. Hence this query is on two tables. Due to predicates
in the WHERE clause, the joins between {\tt t0}, {\tt t1}, and {\tt t3} are
in fact self-joins, and with {\tt t2}, an equi-join. The optimizer generates
different plans for this query as the estimated cardinality of the {\tt
  ft\_HT1} table varies.
\shorten{
Assume initially that the cardinalities (actual cardinalities for the simple
query as an isolated query, estimated cardinalities when the query is an
excerpt) are entirely accurate. In this case, the cardinalities of the two
tables named in the FROM clause were 1,000,000 tuples. With this information,
the cost-based query optimizer presumably generates the best plan for this
query. We'll call this the {\em actual plan}, or simply Plan~0.

However, perhaps the cardinality estimates are inaccurate. Perhaps they do
not reflect the number of tuples in a base or intermediate table. With
this inaccurate information, the cost-based query optimizer generates
perhaps the same plan (that would be nice!) or, more likely, a somewhat less
efficient plan when evaluated on the {\em actual} base or intermediate
table. In this particular experiment, all the cardinalities are accurate
except for one of the tables in the from clause, \verb.ft_HT1.. Let's vary the
inaccuracy from exact to increasingly inaccurate by overriding the
cardinality statistics for \verb.ft_HT1. without changing the physical table. We
then examine the sequence of increasingly poorer plans that result.

One would expect that this sequence of plans is stable. Specifically, for
minor inaccuracy, one would expect that the same plan, Plan~0, to be
produced. And indeed, as the cardinality estimate becomes less and less accurate,
ranging from 570,000 tuples to 1,230,000 tuples, the optimizer continues with
Plan~0. At some point, the
inaccuracy passes a threshold and the optimizer chooses a different plan,
Plan~1 at 1,240,000 tuples and Plan~2 at 560,000 tuples.
Later we hit another threshold (here, at 480,000 tuples), and enter into
a new regime, where Plan~3 is chosen.

For some queries, we see an interesting phenomenon. For these queries
(generally different queries for each DBMS), as we increase the inaccuracy,
the optimizer returns to an {\em earlier} plan! Sometimes, in fact, the
optimizer starts oscillating between two plans, sometimes even switching
back and forth when the cardinality estimate changes by a small percentage.
We had no idea that such strange behavior would occur until we came across
it while testing our experiment manager. We call this phenomenon, in which
the query optimizer returns to a previous plan as the inaccuracy increases,
``query optimizer flutter,'' or simply ``flutter.''}  In fact, for this
query, the query optimizer returns to a previous plan as the inaccuracy
increases, which we call ``query optimizer flutter,'' or simply ``flutter.''

\begin{figure*}[tb]
\centering
%originally 30pc
\epsfig{figure=figures/plan769.eps,width=36pc}
\caption{Suboptimality and Fluttering in Query 769\label{fig:query769}}
\end{figure*}

\shorten{This behavior occurs in Query~769, as shown in}In Figure~\ref{fig:query769},
\shorten{which we term a {\em flutter diagram}. Here,} the \hbox{$x$-axis} depicts the
estimated cardinality and the $y$-axis the execution time of the plan on the
actual cardinality, which is 1,000,000 tuples, denoted by $x=10$. So for example Plan~3 is
selected when the optimizer {\em thought} that the cardinality of table {\tt
  ft\_HT1} was 400K rows; when that plan was run on a table with 1M rows, it
was twice as fast as Plan~0, the plan the optimizer selected when the
optimizer had an accurate cardinality estimate of 1M rows.

\shorten{In this flutter diagram, w}We see that as the cardinality is {\em
  over-}estimated, the optimizer switches to Plan~1 (which in fact is better than Plan~0) 
and stays reliably with that plan. As the cardinality 
is {\em under-}estimated, the optimizer switches to a better one, 
Plan~2 (as this plan is in fact faster than Plan~0, when run on a table with
1,000,000 tuples), and then to an even better one, Plan~3, and eventually,
reaches a sub-optimal plan, Plan~4. 
It is as if the optimizer can't make up its mind. This
behavior is somewhat unnerving.

As we will see shortly, flutter and suboptimality are all around us: {\em
  every} \hbox{DBMS} that we have examined, including several open source
\hbox{DBMSes} and several proprietary \shorten{commercial} \hbox{DBMSes},
covering much of the installed base worldwide, exhibit these phenomena, even
when optimizing very simple queries. Hence, such phenomena are not dependent
on a particular implementation of cost-based optimization. Rather, they seem
to be common to {\em any} cost-based optimizer, independent of the specific
cardinality estimation or plan costing or plan enumeration algorithm or
implementation.

This is where the shift to a new methodology comes in. We want to understand
cost-based optimization deeply, in this case to determine if a fundamental
limitation exists, which means that we need to study multiple instances of
that optimization architecture. To do so, we will articulate a model for how
and in what circumstances suboptimality arises and will provide compelling
evidence via hypothesis testing that this model can predict the behavior of
these several \hbox{DBMSes}. This is what we mean by ``empirical
generalization'' and why it is needed to answer such questions.\shorten{ In
  the long history of research in database query optimization, or even of
  databases in general, our model and its hypothesis tests are the first
  predictive results that we are aware of that apply {\em across}
  \hbox{DBMSes}, rather than on a single, specific \hbox{DBMS} or on a
  specific algorithm. This is due to the adoption of a new research
  methodology, that of empirical generalization.}