### 2017-08-08 subopt-tods17/OLDtex/onepass.tex

\subsection{{\tt OnePass} Scenario}\label{sec:onepass}

In {\tt OnePass} Scenario, a query is studied for suboptimality by keeping track 
of the {\em states} on the variable-table. 
In other words, we locate different query plans at adjacent cardinalities 
defined by the {\em even} or {\em odd} states, 
and query executions are made at both of the states simultaneously, 
while decreasing the variable-table cardinality in one pass.

Indeed, the adjacent cardinalities are not really called even and odd, 
but they are next to each other. 
They, thus, are simply separated by whether a given cardinality over 
granularity is even or odd. 
For instance, the even state will have 2M, 1.98M, 1.96M, $\cdots$, 1M cardinalities 
in order, while the odd state will take 1.99M, 1.97M, $\cdots$, 1.01M cardinalities 
(decremented by granularity (10K) from the even state's cardinality).

Each state initializes its table(s) same as the variable-table(s). 
In addition, it has the {\em pre-compiled} query plan at its own cardinality 
so that the given query can be executed by the {\em prepared} plan at the cardinality 
without any compiling overhead.

While stepping down from the max to the min cardinality, 
if different query plans are observed at the even and odd states, 
then the prepared query execution at both states is made at the same time 
unless the query has been executed at the cardinality before. 
In this regard, for repeatability we time the same query up to 10 times. 
The timing results at each state are inserted into the labshelf for 
suboptimality analysis. 
For more details about timing single query execution, 
please refer to Section~\ref{sec:timing}.

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} analyzeQuery($q$, $varTables$):
\STATE $evenState$.$initialize$($varTables$) \\
\STATE $oddState$.$initialize$($varTables$) \\
\FOR{$card$ = 2M to 1M by 10K}
	\IF{$\dfrac{card}{10K}$ is $even$}
		\STATE $evenState$.$card$ $\leftarrow$ $card$ \\
    	\STATE $evenState$.$plan$ $\leftarrow$ $evenState$.$getPreparedQueryPlan$($q$) \\
    \ELSE
 	    \STATE $oddState$.$card$ $\leftarrow$ $card$ \\
	    \STATE $oddState$.$plan$ $\leftarrow$ $oddState$.$getPreparedQueryPlan$($q$) \\
    \ENDIF
    \IF{($evenState$.$plan$ $\neq$ $oddState$.$plan$)}
		\IF{$\neg${$evenState$}.$isExecuted$()}
			\FOR{$i$ $\leftarrow$ 1 {\bf to} 10}
				\STATE $timeSingleQueryExecution$($i$, $q$,  \\
						\hspace{14.0mm}$evenState$.$plan$, $evenState$.$card$) \\
			\ENDFOR
		\ENDIF
		\IF{$\neg${$oddState$}.$isExecuted$()}
			\FOR{$i$ $\leftarrow$ 1 {\bf to} 10}
				\STATE $timeSingleQueryExecution$($i$, $q$,  \\
						\hspace{14.0mm}$oddState$.$plan$, $oddState$.$card$) \\	
			\ENDFOR
		\ENDIF
	\ENDIF
\ENDFOR 
\end{algorithmic}
\caption{Algorithm for Analyzing a Query by {\tt OnePass} Scenario\label{alg:onepass}}
\end{center}
\end{figure}