### 2017-08-08 subopt-tods17/OLDtex/summary.tex

\section{Summary and Future Work}\label{sec:future}
{\todo{Rick has the modify token}
\todo{Talk about where more work is needed, based on the results}
This paper studies an important
component of a DBMS, the query optimizer. This component is an amazingly
sophisticated piece of code, but is still not well understood after decades
of research and development.

This paper makes the following contributions in
an attempt to gain a new understanding of this component.
\begin{itemize}
\item Introduces a new methodological
perspective, that of {\em empirical generalization}.
\item Uses that perspective to study {\em query suboptimality}, a problem
  not previously investigated.
\item Proposes a {\em predictive model} that attempts a causal explanation of
  suboptimality. 
\item Utilizes a novel {\em research infrastructure}, \azdb, to express, run, and
  analyze experiments.
\item Tests several hypotheses deductively derived from the causal model; in
  each case, the null hypothesis was rejected.
\item Provides a {\em space of hypotheses} that in concert through further
  investigation can refine and support (or not) this model or suggest a
  better model.
\item Thereby provides a path toward scientific progress in the
  understanding of a key enabling technology.
\end{itemize}

We studied a particular
phenomenon, that of the optimizer selecting a wrong plan, indicated by the existence of a query plan that performs more efficiently
than the DBMS' chosen plan, for the same query. From the engineering
perspective, it is of critical importance to understand the prevalence
of suboptimality and its origin(s). Our theory is that suboptimality
is caused in part by the inherent complexity of these system and the
concomitant unanticipated interactions between various rules in the
optimizer.

Through a series of experiments managed by a novel Laboratory
Information Management System termed \azdb, we uncovered compelling evidence 
(a) that suboptimality occurs, (b)~that  suboptimality
correlates with an aspect of query complexity, that of the number of
correlation names used in the query, (c)~that suboptimality correlates
with the cardinality of the effective plan space 
(CEPS)---another (query-specific) measure of optimizer complexity, and (d)~that
query complexity negatively correlates with query complexity and with optimizer
complexity. All of these findings were predicted by our theory.

\comment{
We presented a taxonomy of flutter types and found that the three DBMSes
we studied differed greatly in the types of flutter they produced. We found
that, as expected, flutter was correlated to the number of correlation names 
in the FROM clause. We also found a correlation between the number of 
self-joins and flutter prevalence, and a negative correlation between 
self-joins on the variable table and flutter prevalence. On both, our theory is silent.

Hence, every interaction predicted by our theory was borne out by this experiment, 
thus providing support for our theory.

We also uncovered the surprising result that for many queries, a majority of
the ones we considered, the optimizer picked the wrong query, even when the
cardinality estimates were completely accurate and even for our simplistic
queries. (Note, however, that these results are for only one DBMS.)

One potential engineering ramification of this new understanding is that, 
in cases where the DBMS isn't as sure about the cardinalities of the underlying 
relations or the speed of the disk (if such relations migrated frequently~\cite{reiss03}), the optimizer
shouldn't try as hard to wring out the ``perfect plan,'' Perhaps optimizers
should instead explicitly take uncertainty into account. Indeed, others have
started to argue that uncertainties in the query planning process should be
acknowledged and exploited~\cite{Babcock05}. But those are engineering
considerations; our focus in this work is on the science and on
understanding at a fundamental level these complex systems. Good engineering
should build on solid scientific results.
}
Note that our theory doesn't apply to just one algorithm or to one DBMS.
Rather, it is quite broad, applying to any DBMS with a cost-based
optimizer. We plan to augment \azdb\ to enable other DBMSes to be studied; 
our theory makes specific predictions on what we may observe with these other systems.

\comment{These results also raise several unsettling concerns. These particularly simple
queries used only a few plans within any DBMS. Other more complex queries
will certainly use more plans. These results imply that as a DBMS becomes
more sophisticated, by adding operators, it also become more unstable, as
the prevalence of flutter grows. Perhaps there is a fixed level of
complexity beyond which a cost-based optimizer will be inexorably unstable.}

There are several directions we wish to take this work. 
\c2j{}{Our study raise interesting questions about self-joins, with correlations 
in opposite directions observed. Might this be DBMS-specific? Or is there 
an overarching theory to explain this phenomenon?
}
These investigations are based on very simple SPJ (select-project-join)
queries.  We wish to look at {\em schema complexity}:
the influence of foreign keys and indexes, {\em query complexity}: complex
predicates, subqueries, and aggregates, and {\em instance complexity}:
skewed data.
We also want to manipulate DBMSes internally (at least for
those that are open-source), turning on and off the rules and observing
suboptimality.
%\todo{Rick: DeWitt: anecdotal, broken: first scientific evidence}
Might there be other unanticipated interactions,
that are not known about
simply because they haven't been looked for?

A second concern is the prevalence of sub-optimal plans uncovered in this
work. 
We plan to examine sub-optimality within the other
DBMSes. We need to look at the three complexities listed above also within
this investigation scenario. Does sub-optimality increase with complexity?
Is there a complexity whereby sub-optimality is the rule?

Our predictive model postulates two causal factors to suboptimality:
optimizer complexity and the cost model, the latter specifically, non-linearity of
the cost formul\ae\ and any cost model inaccuracies. In this paper we
considered portions of both factors, but showed only that optimizer
complexity definitely plays a role. It is important to also evaluate the
contribution of non-linearity and cost model inaccuracies. For some DBMSes,
the result of {\tt EXPLAIN PLAN} includes estimated cost of the plan, or even of
each stage of the plan. For these systems, that cost (estimate) is a
dependent variable, one that can be observed. If multiple plans include the
same operator, it may be possible to assemble, from outside the DBMS, an
approximation of the cost formula for that operator, and thus directly
observe whether it is non-linear and/or inaccurate, and then to understand
the impact of that dependent variable on suboptimality. In doing so, one
could test additional hypotheses derived from the causal model introduced in
this paper, ascertain the relative contributions of these two central
factors of the model to suboptimality, and perhaps also identify other
general factors that contribute to suboptimality across DBMSes. The result
would be a truly general, predictive, causal model of this vital component
of a DBMS.