### 2017-10-12 mlppi-ksii17/review/cover-letter.tex

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
\usepackage{color}
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
\hfill School of Computer Science and Engineering
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Kyungpook National University 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Daegu, 41566 Republic of Korea, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-53-950-6372
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: yksuh@knu.ac.kr
\end{flushright} 
%\vfill % forces letterhead to top of page

%\vspace{-0.5in}

\opening{Dear Dr. Mohammad Shojafar (Editor)}

\noindent We are submitting our revised manuscript (ID: TIIS-IS-2017-Jul-0844), entitled ``MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads,'' to KSII Transactions on Internet and Information Systems (TIIS). For this revision we carefully addressed each reviewer's comments. More details are found in the response letter included in this submission. Besides, our submission includes this cover letter and the revised PDF of our paper. Thank you for your support.

\closing{Sincerely yours,}

\end{letter}

\end{document}