### 2018-09-28 pt_characterization/machine_cons_check/daemon_analysis/paper.tex
Rick's request

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Daemon Analysis across Different Machines}

\author{
Young-Kyoon Suh\\
Data \& Knowledge Engineering Laboratory\\
School of Computer Science and Engineering\\
Kyungpook National University\\
}
\maketitle

\section{Investigation on Some Daemons' Appearance Across Machines}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC1 &  1,000 & kslowd000 & sodb8 & 1 & 1 & 1 & 19 \\ \hline
 &   &  & sodb9 & 1 & 1 & 1 & 15 \\ \hline
 &   &  & sodb10 & 1 & 1 & 1 & 19 \\ \hline 
 \hline
  &  & kslowd001 & sodb8 & 1 & 1 & 1 & 15 \\ \hline
 &   &  & sodb9 & 1 & 1 & 1 & 18 \\ \hline
 &   &  & sodb10 & 1 & 1 & 1 & 15 \\ \hline 
 %&   &  & sodb12 & 1 & 1 & 1 & 19 \\ \hline 
%  \hline
 \end{tabular}
  }
 \caption{Daemon statistics on the INC1 data across machines~\label{tab:inc1}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 MAT8K & 300 & grep & sodb12 & 1 & 1 & 1 & 4 \\ \hline
 \hline
  &  & kslowd000  & sodb12 & 16540 & 17033 & 16780.43 & 300 \\ \hline
 &  & kslowd000	& sodb10 & 102 & 107 & 103.33 & 300 \\ \hline 
 \hline
  &  & kslowd001  & sodb12 & 16541 & 17000 & 16777.18 & 300 \\ \hline
  &  & kslowd001	& sodb10 & 102 & 108 & 103.27 & 300 \\ \hline 
 \hline
  &  & sshd  & sodb12 & 1 & 20 & 5.91 & 43 \\ \hline
 &  & sshd & sodb10 & 1 & 8  &  4.17 & 42 \\ \hline 
%   \hline
 \end{tabular}
  }
 \caption{Daemon statistics on the MAT8K data across machines~\label{tab:mat8K}}
\end{table}

\end{document}