### 2018-09-29 pt_characterization/machine_cons_check/daemon_analysis/paper.tex
Daemon analysis done

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Daemon Analysis across Different Machines}

\author{
Young-Kyoon Suh\\
Data \& Knowledge Engineering Laboratory\\
School of Computer Science and Engineering\\
Kyungpook National University\\
}
\maketitle

\section{Investigation on Some Daemons' Appearance Across Machines}

\subsection{INC1, INC16, and INC128}

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
Machine & Task Length & Description & Time Length\\ \hline
{\tt sodb8} (plugged into the {\em bottom left} power strip) & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
{\tt sodb9}  (plugged into the {\em upper left} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
{\tt sodb10}  (plugged into the {\em upper left} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
{\tt sodb12}  (plugged into the {\em upper right} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Detailed description of the INC1, INC16, INC128 data used for histograms\label{tab:exp_notesINCs}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC1 &  1,000 & {\tt kslowd000} & {\tt sodb8} & 1 & 1 & 1 & 19 \\ \hline
 &   &  & {\tt sodb9} & 1 & 1 & 1 & 15 \\ \hline
 &   &  & {\tt sodb10} & 1 & 1 & 1 & 19 \\ \hline 
 \hline
  &  & {\tt kslowd001} & {\tt sodb8} & 1 & 1 & 1 & 15 \\ \hline
 &   &  & {\tt sodb9} & 1 & 1 & 1 & 18 \\ \hline
 &   &  & {\tt sodb10} & 1 & 1 & 1 & 15 \\ \hline 
 %&   &  & {\tt sodb12} & 1 & 1 & 1 & 19 \\ \hline 
%  \hline
 \end{tabular}
  }
 \caption{\textbf{The statistics of the daemons} on the INC1 data across machines~\label{tab:inc1}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC16 &  1,000 & {\tt kslowd000} & {\tt sodb8} & 1 & 1 & 1 & 243 \\ \hline
 &   &  & {\tt sodb9} & 1 & 2 & 1.01 & 242 \\ \hline
 &   &  & {\tt sodb10} & 1 & 1 & 1 & 241 \\ \hline 
 \hline
  &  & {\tt kslowd001} & {\tt sodb8} & 1 & 1 & 1 & 243 \\ \hline
 &   &  & {\tt sodb9} & 1 & 2 & 1 & 241 \\ \hline
 &   &  & {\tt sodb10} & 1 & 1 & 1 & 243 \\ \hline 
 %&   &  & {\tt sodb12} & 1 & 1 & 1 & 19 \\ \hline 
%  \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the INC16 data across machines~\label{tab:inc16}}
\end{table}

\begin{table}[H]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC128 &  300 & {\tt kslowd000} & {\tt sodb8} & 6 & 8 & 6.43 & 300 \\ \hline
 &   &  & {\tt sodb9} & 6 & 7 & 6.42 & 300 \\ \hline
 &   &  & {\tt sodb10} & 6 & 7 & 6.42 & 300 \\ \hline 
 \hline
  &  & {\tt kslowd001} & {\tt sodb8} & 6 & 7 & 6.43 & 300 \\ \hline
 &   &  & {\tt sodb9} & 6 & 8 & 6.42 & 300 \\ \hline
 &   &  & {\tt sodb10} & 6 & 8 & 6.43 & 300 \\ \hline 
 %&   &  & {\tt sodb12} & 1 & 1 & 1 & 19 \\ \hline 
%  \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the INC128 data across machines~\label{tab:inc128}}
\end{table}

\clearpage
\newpage
\subsection{INC1024}

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
Machine & Task Length & Description & Time Length\\ \hline
{\tt sodb9} (plugged into {\em the upper left} power strip) & INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21  \\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Detailed description of the INC1024 data \label{tab:exp_notes1024}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC1024 &  300 & {\tt grep} & {\tt sodb9} & 1 & 7 & 3.23 & 13 \\ \hline
 \hline
  &   & {\tt kslowd000} &  & 51 & 53 & 51.37 & 300 \\ \hline
 \hline
  &  & {\tt kslowd001} &  & 51 & 53 & 51.35 & 300 \\ \hline
 \hline
  &  & {\tt sshd} & & 1 & 18 & 7.55 & 20 \\ \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the INC1024 data across machines~\label{tab:inc1024}}
\end{table}

\clearpage
\newpage
\subsection{INC16384}

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
Machine & Task Length & Description & Time Length\\ \hline
{\tt sodb12} (plugged into {\em the upper right} power strip) & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Detailed description of the INC16384 data \label{tab:exp_notes16384}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC16384 & 130 & {\tt grep} & {\tt sodb12} & 1 & 9 & 3.21& 105 \\ \hline
 \hline
  &  & {\tt sshd} & {\tt sodb12} & 1 & 20 & 8.11 & 140 \\ \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the INC16384 data across machines~\label{tab:inc16384}}
\end{table}

\clearpage
\newpage
\subsection{SORT100, SORT800, MAT1K, and MAT8K}

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{4cm}|p{4cm}|p{4cm}|p{4cm}|} \hline
Machine & Task Length & Description & Time Length\\ \hline
{\tt sodb9}  (plugged into {\em the upper left} power strip) & SORT100, SORT800 & Runs of 300 samples & 2018-04-09 $\sim$2018-04-12\\ \hline
%{\tt sodb10} (plugged into {\em the upper left} power strip)  & MATC1K, MATC8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
{\tt sodb10} (plugged into {\em the upper left} power strip)  & MAT1K, MAT8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
%{\tt sodb12} (plugged into {\em the upper right} power strip) & MATR1K, MATR8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
{\tt sodb12} (plugged into {\em the upper right} power strip) & MAT1K, MAT8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Detailed description of the SORT, MAT data\label{tab:exp_notes_sort_mat}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 SORT100 & 300 & {\tt kslowd000}  & {\tt sodb9} & 1 & 1 & 1 & 64 \\ \hline
  &  & {\tt kslowd001}  & 	& 1 & 2 & 1.02 & 63  \\ \hline
  &  & {\tt sshd}  & & 2 & 8 & 5 & 4 \\ \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the SORT100 data across machines~\label{tab:sort100}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 SORT800 & 300 & {\tt kslowd000}  & {\tt sodb9} & 13 & 15 & 13.78 & 300 \\ \hline
  &  & {\tt kslowd001}  & 	& 13 & 15 & 13.78 & 300  \\ \hline
  &  & {\tt sshd}  & & 1 & 9 & 4.8 & 5 \\ \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the SORT800 data across machines~\label{tab:sort800}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 MAT1K & 300 & {\tt kslowd000}  & {\tt sodb12} & 38 & 161 &  150.85 & 52 \\ \hline
 			 &  	    & {\tt kslowd000}	 & {\tt sodb10} & 1 & 1 & 1 & 50 \\ \hline 
 \hline
  &  & {\tt kslowd001}  & {\tt sodb12} & 146 & 161 & 154.28 & 50 \\ \hline
  &  & {\tt kslowd001}	& {\tt sodb10} & 1 & 1 & 1 & 52 \\ \hline 
 \hline
  &  & {\tt sshd}  & {\tt sodb12} & 1 & 7 & 4.25 & 4 \\ \hline
 &  & {\tt sshd} & {\tt sodb10} & 1 & 7  &  4 & 4 \\ \hline 
%   \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the MAT1K data across machines~\label{tab:mat1K}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 MAT8K & 300 & {\tt grep} & {\tt sodb12} & 1 & 1 & 1 & 4 \\ \hline
 \hline
  &  & {\tt kslowd000}  & {\tt sodb12} & 16540 & 17033 & 16780.43 & 300 \\ \hline
 &  & {\tt kslowd000}	& {\tt sodb10} & 102 & 107 & 103.33 & 300 \\ \hline 
 \hline
  &  & {\tt kslowd001}  & {\tt sodb12} & 16541 & 17000 & 16777.18 & 300 \\ \hline
  &  & {\tt kslowd001}	& {\tt sodb10} & 102 & 108 & 103.27 & 300 \\ \hline 
 \hline
  &  & {\tt sshd}  & {\tt sodb12} & 1 & 20 & 5.91 & 43 \\ \hline
 &  & {\tt sshd} & {\tt sodb10} & 1 & 8  &  4.17 & 42 \\ \hline 
%   \hline
 \end{tabular}
  }
 \caption{The statistics of the daemons on the MAT8K data across machines~\label{tab:mat8K}}
\end{table}

\end{document}